---
title: "Timeline Analysis Report"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
    df_print: paged
params:
  dfFlagged: NULL
  lCharts: NULL
  dfSummary: NULL
  dfFlaggedGroups: NULL
  strStudyID: NULL
  strGroupLevel: "Site"
  strReportDate: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.width = 10,
  fig.height = 6
)
library(plotly)
library(dplyr)
```

# Overview {.tabset}

```{r header-info}
study_text <- if (!is.null(params$strStudyID)) {
  paste0("**Study:** ", params$strStudyID, "
")
} else {
  ""
}
```

`r study_text`
**Report Date:** `r params$strReportDate`

**Group Level:** `r params$strGroupLevel`

***

## Summary Statistics

```{r summary-table}
if (!is.null(params$dfSummary)) {
  knitr::kable(
    params$dfSummary,
    col.names = c("Statistic", "Value"),
    align = c("l", "r")
  )
}
```

## Flag Distribution

```{r flag-distribution}
if (!is.null(params$dfFlagged)) {
  flag_labels <- c(
    "-2" = "Red Low",
    "-1" = "Amber Low",
    "0" = "Normal",
    "1" = "Amber High",
    "2" = "Red High"
  )
  
  flag_colors <- c(
    "Red Low" = "#D73027",
    "Amber Low" = "#FC8D59",
    "Normal" = "#999999",
    "Amber High" = "#91BFDB",
    "Red High" = "#4575B4"
  )
  
  dfFlagDist <- params$dfFlagged %>%
    mutate(FlagLabel = flag_labels[as.character(Flag)]) %>%
    count(FlagLabel) %>%
    mutate(
      FlagLabel = factor(FlagLabel, levels = c("Red Low", "Amber Low", "Normal", "Amber High", "Red High")),
      Percentage = round(n / sum(n) * 100, 1)
    )
  
  plot_ly(
    dfFlagDist,
    x = ~FlagLabel,
    y = ~n,
    type = "bar",
    marker = list(color = flag_colors[as.character(dfFlagDist$FlagLabel)]),
    text = ~paste0(n, " (", Percentage, "%)"),
    textposition = "auto",
    hoverinfo = "text",
    hovertext = ~paste0(FlagLabel, ": ", n, " observations (", Percentage, "%)")
  ) %>%
    layout(
      title = "Flag Distribution (All Time Points)",
      xaxis = list(title = "Flag Category"),
      yaxis = list(title = "Count"),
      showlegend = FALSE
    )
}
```

***

# Timeline Visualization {.tabset}

## Line Chart

The line chart shows the cumulative metric (Numerator / Denominator) for each `r params$strGroupLevel` over time. Points are colored by their flag status.

```{r timeline-chart, fig.height=7}
if (!is.null(params$lCharts$Timeline_Timeline)) {
  params$lCharts$Timeline_Timeline
}
```

## Boxplot

The boxplot shows the distribution of metric values across all `r params$strGroupLevel`s for each month. Individual points are colored by their flag status.

```{r boxplot-chart, fig.height=7}
if (!is.null(params$lCharts$Timeline_Boxplot)) {
  params$lCharts$Timeline_Boxplot
}
```

***

# Flagged `r params$strGroupLevel`s

The table below shows `r params$strGroupLevel`s that are currently flagged (at the most recent time point).

```{r flagged-table}
if (!is.null(params$dfFlaggedGroups) && nrow(params$dfFlaggedGroups) > 0) {
  dfDisplay <- params$dfFlaggedGroups %>%
    mutate(
      Metric = round(Metric, 4),
      Score = round(Score, 2)
    ) %>%
    select(
      GroupID,
      Month = NMonth,
      Numerator,
      Denominator,
      Metric,
      `Z-Score` = Score,
      Flag = FlagLabel
    )
  
  names(dfDisplay)[1] <- params$strGroupLevel
  
  if (requireNamespace("DT", quietly = TRUE)) {
    DT::datatable(
      dfDisplay,
      options = list(
        pageLength = 25,
        dom = "frtip",
        order = list(list(5, "desc"))
      ),
      rownames = FALSE
    )
  } else {
    knitr::kable(dfDisplay)
  }
} else {
  cat("No flagged", params$strGroupLevel, "s at the current time point.")
}
```

***

# Data Details

## Raw Data Preview

```{r data-preview}
if (!is.null(params$dfFlagged)) {
  dfPreview <- params$dfFlagged %>%
    mutate(
      Metric = round(Metric, 4),
      Score = round(Score, 2)
    ) %>%
    arrange(GroupID, NMonth) %>%
    head(100)
  
  if (requireNamespace("DT", quietly = TRUE)) {
    DT::datatable(
      dfPreview,
      options = list(
        pageLength = 10,
        scrollX = TRUE,
        dom = "frtip"
      ),
      rownames = FALSE,
      caption = "First 100 rows of timeline data"
    )
  } else {
    knitr::kable(head(dfPreview, 20), caption = "First 20 rows of timeline data")
  }
}
```

***

<div style="text-align: center; color: #666; font-size: 0.9em; margin-top: 30px;">
Generated by gsm.timez version `r packageVersion("gsm.timez")`
</div>
