---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# gsm.timez

<!-- badges: start -->
[![R-CMD-check](https://github.com/IMPALA-Consortium/gsm.timez/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/IMPALA-Consortium/gsm.timez/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/IMPALA-Consortium/gsm.timez/graph/badge.svg)](https://app.codecov.io/gh/IMPALA-Consortium/gsm.timez)
<!-- badges: end -->

## Overview

The `gsm.timez` package provides functions for generating cumulative event timelines 
and calculating z-scores for clinical trial monitoring. It is part of the IMPALA 
Consortium's GSM (Good Statistical Monitoring) framework.

The package includes five main functions:

- **`Timeline()`**: Generates a site-level timeline of cumulative numerator 
  events (e.g., adverse events) over sequential months.
- **`TimeZScore()`**: Calculates z-scores for each group (e.g., site) based on 
  the rate of events, enabling identification of sites with unusual event patterns.
- **`Flag()`**: Adds flag columns to analyzed data identifying possible statistical 
  outliers based on threshold comparisons. This is an alias for `gsm.core::Flag()`.
- **`Visualize()`**: Creates a line plot showing site metrics over time with dots 
  colored by flag value to highlight outliers.
- **`VisualizeBoxplot()`**: Creates boxplots showing the distribution of Metric 
  values for each month, with individual points colored by flag value.

## Installation

You can install the development version of gsm.timez from [GitHub](https://github.com/) with:
``` r
# install.packages("pak")
pak::pak("IMPALA-Consortium/gsm.timez")
```

## Quick Start

```{r quick-start, message=FALSE, warning=FALSE}
library(gsm.timez)
library(dplyr)

# Prepare data
dfSubjects <- clindata::rawplus_dm
dfNumerator <- clindata::rawplus_ae
dfDenominator <- clindata::rawplus_visdt %>%
  dplyr::mutate(visit_dt = as.Date(visit_dt, "%Y-%m-%d"))

# Full pipeline: Timeline -> Z-Score -> Flag -> Visualize
gsm.timez::Timeline(
  dfSubjects            = dfSubjects,
  dfNumerator           = dfNumerator,
  dfDenominator         = dfDenominator,
  strGroupCol           = "siteid",
  strSubjectCol         = "subjid",
  strNumeratorDateCol   = "aest_dt",
  strDenominatorDateCol = "visit_dt"
) %>%
  gsm.timez::TimeZScore() %>%
  gsm.timez::Flag() %>%
  gsm.timez::Visualize()
```

## Learn More

For detailed documentation on each function, including parameters, output columns, 
and advanced usage examples, see the [Detailed Usage vignette](https://impala-consortium.github.io/gsm.timez/articles/DetailedUsage.html).

## Quality Control

Since {gsm} is designed for use in a GCP framework, we have conducted extensive quality control as part of our development process. In particular, we do the following during early development:

- **Unit Tests** - Unit tests are written for all core functions, 100% coverage required.
- **Workflow Tests** - Additional unit tests confirm that core workflows behave as expected.
- **Function Documentation** - Detailed documentation for each exported function with examples is maintained with Roxygen.
- **Package Checks** - Standard package checks are run using GitHub Actions and must be passing before PRs are merged.
- **Continuous Integration** - Continuous integration is provided via GitHub Actions.
- **Code Formatting** - Code is formatted with {styler} before each release.
- **Contributor Guidelines** - Detailed contributor guidelines including step-by-step processes for code development and releases are provided as a vignette.
- **Code Demonstration** - Cookbook Vignette provides demos and explanations for code usage.
