---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dpi = 150
)
```

# gsm.timez

<!-- badges: start -->
[![R-CMD-check](https://github.com/IMPALA-Consortium/gsm.timez/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/IMPALA-Consortium/gsm.timez/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/IMPALA-Consortium/gsm.timez/graph/badge.svg)](https://app.codecov.io/gh/IMPALA-Consortium/gsm.timez)
<!-- badges: end -->

## Overview

`gsm.timez` is a GSM extension for longitudinal site monitoring. It applies funnel
plot scoring across sequential months to detect sites with unusual cumulative event
trajectories.

The package includes the following main functions:

- **`Timeline()`**: Generates a site-level timeline of cumulative numerator
  events (e.g. adverse events) over sequential months.
- **`Analyze_TimeZFunnel()`**: Applies `gsm.core::Analyze_NormalApprox()` across
  sequential months, producing funnel plot scores for each site over time.
- **`Flag()`**: Extends `gsm.core::Flag()` with sparse-month awareness for
  time-series data.
- **`PredictBounds_TimeZFunnel()`**: Applies `gsm.core::Analyze_NormalApprox_PredictBounds()`
  across sequential months to produce funnel bounds for visualization.
- **`Visualize_Funnel()`**: funnel plot for a single selected month, showing
  site Metric values against their Denominator with Normal approximation control bounds.
- **`Visualize_Heatmap()`**: heatmap showing each site's flag status over time.
- **`Visualize_Site()`**:  line plot highlighting a single site's cumulative
  metric trajectory against all other sites, marking flagged months.

## Installation

You can install the development version of gsm.timez with:
```{r, eval=FALSE}
install.packages("pak")
pak::pak("IMPALA-Consortium/gsm.timez")
```

## Quick Start

```{r quick-start, message=FALSE, warning=FALSE, eval=FALSE}
library(gsm.timez)
library(dplyr)

# Prepare data
dfSubjects <- clindata::rawplus_dm
dfNumerator <- clindata::rawplus_ae
dfDenominator <- clindata::rawplus_visdt %>%
  dplyr::mutate(visit_dt = as.Date(visit_dt, "%Y-%m-%d"))

# Full pipeline: Timeline -> Funnel Score -> Flag -> Visualize
gsm.timez::Timeline(
  dfSubjects            = dfSubjects,
  dfNumerator           = dfNumerator,
  dfDenominator         = dfDenominator,
  strGroupCol           = "invid",
  strSubjectCol         = "subjid",
  strNumeratorDateCol   = "aest_dt",
  strDenominatorDateCol = "visit_dt"
) %>%
  gsm.timez::Analyze_TimeZFunnel() %>%
  gsm.timez::Flag(vThreshold = c(-1.5, -1, 2, 3)) %>%
  gsm.timez::Visualize_Heatmap(nSites = 30)
```

```{r quick-start-plot, message=FALSE, warning=FALSE, echo=FALSE, dev.args=list(bg="transparent")}
library(gsm.timez)
library(dplyr)

dfSubjects <- clindata::rawplus_dm
dfNumerator <- clindata::rawplus_ae
dfDenominator <- clindata::rawplus_visdt %>%
  dplyr::mutate(visit_dt = as.Date(visit_dt, "%Y-%m-%d"))

gsm.timez::Timeline(
  dfSubjects            = dfSubjects,
  dfNumerator           = dfNumerator,
  dfDenominator         = dfDenominator,
  strGroupCol           = "invid",
  strSubjectCol         = "subjid",
  strNumeratorDateCol   = "aest_dt",
  strDenominatorDateCol = "visit_dt"
) %>%
  gsm.timez::Analyze_TimeZFunnel() %>%
  gsm.timez::Flag(vThreshold = c(-1.5, -1, 2, 3)) %>%
  gsm.timez::Visualize_Heatmap(nSites = 30) +
  ggplot2::theme(plot.background = ggplot2::element_rect(fill = "transparent", colour = NA))
```

## Quality Control

Since {gsm} is designed for use in a GCP framework, we have conducted extensive quality control as part of our development process. In particular, we do the following during early development:

- **Unit Tests** - Unit tests are written for all core functions, 100% coverage required.
- **Workflow Tests** - Additional unit tests confirm that core workflows behave as expected.
- **Function Documentation** - Detailed documentation for each exported function with examples is maintained with Roxygen.
- **Package Checks** - Standard package checks are run using GitHub Actions and must be passing before PRs are merged.
- **Continuous Integration** - Continuous integration is provided via GitHub Actions.
- **Code Formatting** - Code is formatted with {styler} before each release.
- **Contributor Guidelines** - Detailed contributor guidelines including step-by-step processes for code development and releases are provided as a vignette.
- **Code Demonstration** - [Cookbook Vignette](https://impala-consortium.github.io/gsm.timez/articles/Cookbook.html) provides demos and explanations for code usage.
