---
title: "Cookbook"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cookbook}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`gsm.timez` extends the GSM ecosystem with longitudinal site-level monitoring.
Where standard GSM metrics provide a cross-sectional snapshot of each site's
performance, `gsm.timez` tracks how cumulative event rates evolve month by
month, applying `gsm.core::Analyze_NormalApprox()` at each time step to flag
sites whose trajectory diverges from the study-wide trend.

This vignette provides an overview of the main functions in `gsm.timez` with
usage examples.

## Setup

```{r setup, message=FALSE}
library(gsm.timez)
library(dplyr)
```

## Data Preparation

The examples use `clindata` sample data. Install it from GitHub
([Gilead-BioStats/clindata](https://github.com/Gilead-BioStats/clindata)) to
explore the input datasets (`rawplus_dm`, `rawplus_ae`, `rawplus_visdt`) and
understand the expected column structure.

```{r data-prep}
dfSubjects <- clindata::rawplus_dm

dfNumerator <- clindata::rawplus_ae

dfDenominator <- clindata::rawplus_visdt %>%
  dplyr::mutate(visit_dt = as.Date(visit_dt, "%Y-%m-%d"))
```

## Complete Pipeline

The full analysis chain from raw data to visualization:

```{r complete-pipeline}
gsm.timez::Timeline(
  dfSubjects            = dfSubjects,
  dfNumerator           = dfNumerator,
  dfDenominator         = dfDenominator,
  strGroupCol           = "invid",
  strSubjectCol         = "subjid",
  strNumeratorDateCol   = "aest_dt",
  strDenominatorDateCol = "visit_dt"
) %>%
  gsm.timez::Analyze_TimeZFunnel() %>%
  gsm.timez::Flag(vThreshold = c(-1.5, -1, 2, 3)) %>%
  gsm.timez::Visualize_Heatmap(nSites = 10)
```

## Step-by-Step Breakdown

The following sections walk through each function in the pipeline individually.

### Timeline

`Timeline()` generates a site-level timeline of cumulative numerator events over
sequential months, producing one row per site-month combination:

```{r timeline-basic}
dfTimeline <- gsm.timez::Timeline(
  dfSubjects            = dfSubjects,
  dfNumerator           = dfNumerator,
  dfDenominator         = dfDenominator,
  strGroupCol           = "invid",
  strSubjectCol         = "subjid",
  strNumeratorDateCol   = "aest_dt",
  strDenominatorDateCol = "visit_dt"
)

knitr::kable(head(dfTimeline))
```

### Analyze_TimeZFunnel

`Analyze_TimeZFunnel()` applies `gsm.core::Analyze_NormalApprox()` at each
monthly cross-section to calculate a funnel plot score for each site:

```{r zscore-basic}
dfTimeZScore <- gsm.timez::Analyze_TimeZFunnel(dfTimeline)
```

### Flag

`Flag()` adds a `Flag` column identifying statistical outliers based on z-score
thresholds:

```{r flag}
dfFlagged <- gsm.timez::Flag(dfTimeZScore, vThreshold = c(-1.5, -1, 2, 3))
```

### Visualize_Heatmap

`Visualize_Heatmap()` creates a heatmap showing each site's flag status over
time:

```{r heatmap-height, include=FALSE}
# Calculate height of the heatmap below so that all sites fit
nSitesHeatmap <- dplyr::n_distinct(dfFlagged$GroupID)
figHeightHeatmap <- nSitesHeatmap * 0.4 + 2
```

```{r visualize-heatmap, fig.width=8, fig.height=figHeightHeatmap}
gsm.timez::Visualize_Heatmap(dfFlagged)
```

### Visualize_Funnel

`Visualize_Funnel()` creates a funnel plot for a selected month, showing site
Metric values against their Denominator. The bounds are computed by
`PredictBounds_TimeZFunnel()`, which applies
`gsm.core::Analyze_NormalApprox_PredictBounds()` at each month:

```{r visualize-funnel, fig.width=8, fig.height=5}
dfBounds <- gsm.timez::PredictBounds_TimeZFunnel(dfTimeZScore, vThreshold = c(-1.5, -1, 2, 3))
gsm.timez::Visualize_Funnel(dfFlagged, dfBounds = dfBounds)
```

### Visualize_Site

`Visualize_Site()` plots a single site's cumulative metric trajectory against
all other sites, with colored dots marking flagged months:

```{r visualize-site, fig.width=8, fig.height=5}
gsm.timez::Visualize_Site(dfFlagged, dfBounds = dfBounds, strSiteID = "0X159")
```

For YAML workflow integration and HTML report generation, see
`vignette("gsm-workflow", package = "gsm.timez")`.
