---
title: "Detailed Usage and Advanced Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Detailed Usage and Advanced Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette provides detailed documentation for each function in `gsm.timez`, 
including parameter descriptions, output columns, and advanced usage examples.

## Setup

```{r setup, message=FALSE}
library(gsm.timez)
library(dplyr)
```

## Data Preparation

We'll use sample clinical trial data from the `clindata` package to demonstrate 
the functions. The key datasets are:
- `rawplus_dm`: Subject demographics with site assignments
- `rawplus_ae`: Adverse event records with dates
- `rawplus_visdt`: Visit date records

```{r data-prep}
# Load subject data
dfSubjects <- clindata::rawplus_dm

# Load adverse events (numerator)
dfNumerator <- clindata::rawplus_ae

# Load visit dates (denominator) - convert to Date type
dfDenominator <- clindata::rawplus_visdt %>%
  dplyr::mutate(visit_dt = as.Date(visit_dt, "%Y-%m-%d"))
```

Let's preview each dataset:

```{r data-preview}
# Subjects: key columns are subjid (subject ID) and invid (site/group)
knitr::kable(head(dfSubjects[, c("subjid", "invid", "studyid")], 5))

# Numerator (AEs): key columns are subjid and aest_dt (AE start date)
knitr::kable(head(dfNumerator[, c("subjid", "aest_dt")], 5))

# Denominator (Visits): key columns are subjid and visit_dt (visit date)
knitr::kable(head(dfDenominator[, c("subjid", "visit_dt")], 5))
```

## Timeline Function

The `Timeline()` function generates a site-level timeline of cumulative 
numerator events over sequential months.

### Parameters

- `dfSubjects`: Data frame with subject IDs and grouping variables
- `dfNumerator`: Data frame with event records (numerator)
- `dfDenominator`: Data frame with time point records (denominator)
- `strGroupCol`: Column name for grouping (e.g., "invid")
- `strSubjectCol`: Column name for subject ID
- `strNumeratorDateCol`: Date column in numerator data
- `strDenominatorDateCol`: Date column in denominator data

### Basic Usage

```{r timeline-basic}
dfTimeline <- gsm.timez::Timeline(
  dfSubjects            = dfSubjects,
  dfNumerator           = dfNumerator,
  dfDenominator         = dfDenominator,
  strGroupCol           = "invid",
  strSubjectCol         = "subjid",
  strNumeratorDateCol   = "aest_dt",
  strDenominatorDateCol = "visit_dt"
)

knitr::kable(head(dfTimeline))
```

### Output Columns

The Timeline output contains one row per site-month combination. Column names follow 
GSM standard PascalCase conventions:

- `GroupID`: Group identifier (e.g., site ID)
- `GroupLevel`: Name of the grouping column used
- `Numerator`: Cumulative count of numerator events (e.g., AEs) at the site
- `Denominator`: Cumulative count of visits at the site
- `DenominatorMonth`: The calendar month (Date, truncated to first of month)
- `NMonth`: Sequential month index for the site (1, 2, 3, ...). This is the 
  site's "study month" - the first month with data is 1, the second is 2, etc.

## Analyze_TimeZFunnel Function

The `Analyze_TimeZFunnel()` function takes the output from `Timeline()` and calculates
funnel plot scores using the Poisson-based normal approximation (Zink et al.).
The score adjusts for sample size, so larger sites are held to tighter standards.

The score formula is:

$$Score = \frac{Metric - OverallMetric}{\sqrt{OverallMetric \times Factor / Denominator}}$$

Where `OverallMetric` is the study-wide pooled rate at each month cross-section.

### Basic Usage

The function is designed to work seamlessly with `Timeline()` output using the
pipe operator:

```{r zscore-basic}
dfTimeZScore <- gsm.timez::Analyze_TimeZFunnel(dfTimeline)

knitr::kable(head(dfTimeZScore))
```

### Output Columns

The `Analyze_TimeZFunnel()` output returns the input data frame with additional columns:

- `Metric`: The ratio of Numerator to Denominator (Numerator/Denominator)
- `OverallMetric`: The study-wide pooled rate at each month
- `Factor`: The factor used in the normal approximation for calculating bounds
- `Score`: The adjusted z-score that accounts for sample size

## Flag Function

The `Flag()` function adds a `Flag` column identifying statistical outliers based 
on z-score thresholds:

```{r flag}
dfFlagged <- gsm.timez::Flag(dfTimeZScore)

knitr::kable(head(dfFlagged))
```

## Visualize_Heatmap Function

The `Visualize_Heatmap()` function creates a heatmap showing each site's flag status
over time:

```{r visualize, fig.width=8, fig.height=5}
gsm.timez::Visualize_Heatmap(dfFlagged)
```

## Visualize_Funnel Function

The `Visualize_Funnel()` function creates a funnel plot for a selected month,
showing site Metric values against their Denominator with Poisson-based control bounds:

```{r visualize-boxplot, fig.width=8, fig.height=5}
dfBounds <- gsm.timez::PredictBounds_TimeZFunnel(dfTimeZScore)
gsm.timez::Visualize_Funnel(dfFlagged, dfBounds = dfBounds)
```

## Advanced Usage

### Using Different Grouping Levels

The `strGroupCol` parameter allows you to analyze data at different levels. 
For example, you could group by country instead of site:

```{r different-grouping}
# Group by country instead of site
dfTimelineByCountry <- gsm.timez::Timeline(
  dfSubjects = dfSubjects,
  dfNumerator = dfNumerator,
  dfDenominator = dfDenominator,
  strGroupCol = "country",
  strSubjectCol = "subjid",
  strNumeratorDateCol = "aest_dt",
  strDenominatorDateCol = "visit_dt"
)
```
### Working with Custom Data Sources

When working with data that doesn't follow the `clindata` structure, ensure your 
data frames contain the required columns:

**dfSubjects** must contain:
- A subject ID column (specified by `strSubjectCol`)
- A grouping column (specified by `strGroupCol`)

**dfNumerator** must contain:
- A subject ID column (same as `strSubjectCol`)
- A date column (specified by `strNumeratorDateCol`)

**dfDenominator** must contain:
- A subject ID column (same as `strSubjectCol`)
- A date column (specified by `strDenominatorDateCol`)

```{r custom-data}
# Example with custom column names
my_subjects <- data.frame(
  patient_id = c("P001", "P002", "P003"),
  site = c("Site_A", "Site_A", "Site_B")
)

my_events <- data.frame(
  patient_id = c("P001", "P001", "P002"),
  event_date = as.Date(c("2024-01-15", "2024-02-20", "2024-01-10"))
)

my_visits <- data.frame(
  patient_id = c("P001", "P002", "P003"),
  visit_date = as.Date(c("2024-01-01", "2024-01-05", "2024-01-08"))
)

# Use custom column names
dfTimeline <- gsm.timez::Timeline(
  dfSubjects            = my_subjects,
  dfNumerator           = my_events,
  dfDenominator         = my_visits,
  strGroupCol           = "site",
  strSubjectCol         = "patient_id",
  strNumeratorDateCol   = "event_date",
  strDenominatorDateCol = "visit_date"
)
```

## Interactive Widgets and HTML Reports

For interactive plotly widgets and HTML report generation, see the dedicated 
vignette: `vignette("HTMLReporting", package = "gsm.timez")`.

Key functions include:

- `Widget()` - Create interactive plotly charts for any visualization function
- `Merge_ChartLists()` - Combine chart lists from multiple Widget calls
- `Report_TimeZFunnel()` - Generate a complete HTML report

## Complete Workflow

The full pipeline from raw data to visualization:

```{r complete-workflow}
gsm.timez::Timeline(
  dfSubjects            = dfSubjects,
  dfNumerator           = dfNumerator,
  dfDenominator         = dfDenominator,
  strGroupCol           = "invid",
  strSubjectCol         = "subjid",
  strNumeratorDateCol   = "aest_dt",
  strDenominatorDateCol = "visit_dt"
) %>%
  gsm.timez::Analyze_TimeZFunnel() %>%
  gsm.timez::Flag() %>%
  gsm.timez::Visualize_Heatmap()
```

For generating HTML reports, see `vignette("HTMLReporting", package = "gsm.timez")`.
