---
title: "Detailed Usage and Advanced Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Detailed Usage and Advanced Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette provides detailed documentation for each function in `gsm.timez`, 
including parameter descriptions, output columns, and advanced usage examples.

## Setup

```{r setup, message=FALSE}
library(gsm.timez)
library(dplyr)
```

## Data Preparation

We'll use sample clinical trial data from the `clindata` package to demonstrate 
the functions. The key datasets are:
- `rawplus_dm`: Subject demographics with site assignments
- `rawplus_ae`: Adverse event records with dates
- `rawplus_visdt`: Visit date records

```{r data-prep}
# Load subject data
dfSubjects <- clindata::rawplus_dm

# Load adverse events (numerator)
dfNumerator <- clindata::rawplus_ae

# Load visit dates (denominator) - convert to Date type
dfDenominator <- clindata::rawplus_visdt %>%
  dplyr::mutate(visit_dt = as.Date(visit_dt, "%Y-%m-%d"))
```

Let's preview each dataset:

```{r data-preview}
# Subjects: key columns are subjid (subject ID) and siteid (site/group)
knitr::kable(head(dfSubjects[, c("subjid", "siteid", "studyid")], 5))

# Numerator (AEs): key columns are subjid and aest_dt (AE start date)
knitr::kable(head(dfNumerator[, c("subjid", "aest_dt")], 5))

# Denominator (Visits): key columns are subjid and visit_dt (visit date)
knitr::kable(head(dfDenominator[, c("subjid", "visit_dt")], 5))
```

## Timeline Function

The `Timeline()` function generates a site-level timeline of cumulative 
numerator events over sequential months.

### Parameters

- `dfSubjects`: Data frame with subject IDs and grouping variables
- `dfNumerator`: Data frame with event records (numerator)
- `dfDenominator`: Data frame with time point records (denominator)
- `strGroupCol`: Column name for grouping (e.g., "siteid")
- `strSubjectCol`: Column name for subject ID
- `strNumeratorDateCol`: Date column in numerator data
- `strDenominatorDateCol`: Date column in denominator data

### Basic Usage

```{r timeline-basic}
dfTimeline <- gsm.timez::Timeline(
  dfSubjects            = dfSubjects,
  dfNumerator           = dfNumerator,
  dfDenominator         = dfDenominator,
  strGroupCol           = "siteid",
  strSubjectCol         = "subjid",
  strNumeratorDateCol   = "aest_dt",
  strDenominatorDateCol = "visit_dt"
)

knitr::kable(head(dfTimeline))
```

### Output Columns

The Timeline output contains one row per site-month combination. Column names follow 
GSM standard PascalCase conventions:

- `GroupID`: Group identifier (e.g., site ID)
- `GroupLevel`: Name of the grouping column used
- `Numerator`: Cumulative count of numerator events (e.g., AEs) at the site
- `Denominator`: Cumulative count of visits at the site
- `DenominatorMonth`: The calendar month (Date, truncated to first of month)
- `NMonth`: Sequential month index for the site (1, 2, 3, ...). This is the 
  site's "study month" - the first month with data is 1, the second is 2, etc.

## TimeZScore Function

The `TimeZScore()` function takes the output from `Timeline()` and calculates 
cumulative z-scores for each row using an expanding window indexed by month.

The z-score formula is:

$$z = \frac{Metric - mean(Metrics)}{sd(Metrics)}$$

Where `Metrics` includes all Metric values from all groups where `NMonth <= current_NMonth`.
This means:

- For month 1: z-scores are calculated using all Metrics from month 1
- For month 2: z-scores are calculated using all Metrics from months 1 and 2
- And so on...

If fewer than 2 Metrics exist in the cumulative window, the Score is 0.

### Basic Usage

The function is designed to work seamlessly with `Timeline()` output using the 
pipe operator:

```{r zscore-basic}
dfTimeZScore <- gsm.timez::TimeZScore(dfTimeline)

knitr::kable(head(dfTimeZScore))
```

### Output Columns

The TimeZScore output returns the input data frame with two additional columns:

- `Metric`: The ratio of Numerator to Denominator (Numerator/Denominator)
- `Score`: The z-score calculated using the cumulative expanding window

## Flag Function

The `Flag()` function adds a `Flag` column identifying statistical outliers based 
on z-score thresholds:

```{r flag}
dfFlagged <- gsm.timez::Flag(dfTimeZScore)

knitr::kable(head(dfFlagged))
```

## Visualize Function

The `Visualize()` function creates a line plot showing each site's metric over time, 
with dots colored by flag value:

```{r visualize, fig.width=8, fig.height=5}
gsm.timez::Visualize(dfFlagged)
```

## VisualizeBoxplot Function

The `VisualizeBoxplot()` function creates a boxplot showing the distribution of 
Metric values for each month, with individual points colored by flag value. 
Points are rendered behind the boxplots with reduced opacity (alpha=0.4):

```{r visualize-boxplot, fig.width=8, fig.height=5}
# Show only first 10 months for clarity
gsm.timez::VisualizeBoxplot(dfFlagged, nMonths = 1:10)
```

## Advanced Usage

### Using Different Grouping Levels

The `strGroupCol` parameter allows you to analyze data at different levels. 
For example, you could group by country instead of site:

```{r different-grouping, eval=FALSE}
# Group by country instead of site
dfTimelineByCountry <- gsm.timez::Timeline(
  dfSubjects = dfSubjects,
  dfNumerator = dfNumerator,
  dfDenominator = dfDenominator,
  strGroupCol = "country",
  strSubjectCol = "subjid",
  strNumeratorDateCol = "aest_dt",
  strDenominatorDateCol = "visit_dt"
)
```

### Integration with GSM Workflow System

`gsm.timez` can be used with the GSM YAML workflow system. Workflow files are 
located in `inst/workflow/2_metrics/`:

```{r workflow-example, eval=FALSE}
library(gsm.core)
library(gsm.mapping)

# Step 1: Load raw data
lRaw <- list(
  Raw_SUBJ = clindata::rawplus_dm,
  Raw_AE = clindata::rawplus_ae,
  Raw_VISIT = clindata::rawplus_visdt
)

# Step 2: Run mapping workflows
mapping_wf <- MakeWorkflowList(
  strPath = system.file("workflow/1_mappings", package = "gsm.mapping")
)
spec <- CombineSpecs(mapping_wf)
lIngest <- Ingest(lRaw, spec)
lMapped <- RunWorkflows(mapping_wf, lIngest)

# Step 3: Run gsm.timez metric workflows
metrics_wf <- MakeWorkflowList(
  strPath = system.file("workflow/2_metrics", package = "gsm.timez")
)
lResults <- RunWorkflows(metrics_wf, lMapped)
```

### Working with Custom Data Sources

When working with data that doesn't follow the `clindata` structure, ensure your 
data frames contain the required columns:

**dfSubjects** must contain:
- A subject ID column (specified by `strSubjectCol`)
- A grouping column (specified by `strGroupCol`)

**dfNumerator** must contain:
- A subject ID column (same as `strSubjectCol`)
- A date column (specified by `strNumeratorDateCol`)

**dfDenominator** must contain:
- A subject ID column (same as `strSubjectCol`)
- A date column (specified by `strDenominatorDateCol`)

```{r custom-data, eval=FALSE}
# Example with custom column names
my_subjects <- data.frame(
  patient_id = c("P001", "P002", "P003"),
  site = c("Site_A", "Site_A", "Site_B")
)

my_events <- data.frame(
  patient_id = c("P001", "P001", "P002"),
  event_date = as.Date(c("2024-01-15", "2024-02-20", "2024-01-10"))
)

my_visits <- data.frame(
  patient_id = c("P001", "P002", "P003"),
  visit_date = as.Date(c("2024-01-01", "2024-01-05", "2024-01-08"))
)

# Use custom column names
dfTimeline <- gsm.timez::Timeline(
  dfSubjects            = my_subjects,
  dfNumerator           = my_events,
  dfDenominator         = my_visits,
  strGroupCol           = "site",
  strSubjectCol         = "patient_id",
  strNumeratorDateCol   = "event_date",
  strDenominatorDateCol = "visit_date"
)
```

## Complete Workflow

The full pipeline from raw data to visualization:

```{r complete-workflow, eval=FALSE}
gsm.timez::Timeline(
  dfSubjects            = dfSubjects,
  dfNumerator           = dfNumerator,
  dfDenominator         = dfDenominator,
  strGroupCol           = "siteid",
  strSubjectCol         = "subjid",
  strNumeratorDateCol   = "aest_dt",
  strDenominatorDateCol = "visit_dt"
) %>%
  gsm.timez::TimeZScore() %>%
  gsm.timez::Flag() %>%
  gsm.timez::Visualize()
```
