---
title: "GSM YAML Report Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GSM YAML Report Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r vignette-setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(gsm.timez)
library(gsm.core)
library(gsm.mapping)
```

This vignette demonstrates how to use the GSM YAML workflow system to generate
an HTML report for the timeline KRI.

## Prepare Source Data

```{r prepare-source, message=FALSE}
lSource <- list(
  Source_STUDY = clindata::ctms_study,
  Source_SITE = clindata::ctms_site,
  Source_SUBJ = clindata::rawplus_dm,
  Source_AE = clindata::rawplus_ae,
  # DATAENT is not used by gsm.timez but is required by gsm.mapping's AE workflow.
  # The studyid column is added because gsm.mapping expects it but clindata has protocolname.
  Source_DATAENT = clindata::edc_data_pages %>%
    dplyr::mutate(studyid = protocolname),
  Source_VISIT = clindata::rawplus_visdt
)
```

## Run Mapping Workflows

Use `Ingest()` to standardize column names based on the workflow specs, then
run the mapping workflows to create the mapped data layer.

```{r run-mapping, message=FALSE, results='hide'}
mapping_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/1_mappings", package = "gsm.mapping"),
  strNames = c("SUBJ", "AE", "VISIT", "SITE", "STUDY", "COUNTRY")
)
mapping_spec <- gsm.mapping::CombineSpecs(mapping_wf)
lRaw <- gsm.mapping::Ingest(lSource, mapping_spec)
lMapped <- gsm.core::RunWorkflows(mapping_wf, lRaw)
```

## Run Analysis and Reporting Workflows

```{r run-metrics, message=FALSE, results='hide'}
# Run gsm.timez 2_metrics workflows
metrics_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/2_metrics", package = "gsm.timez")
)
lAnalyzed <- gsm.core::RunWorkflows(metrics_wf, lMapped)

# Run gsm.timez 3_reporting workflows
reporting_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/3_reporting", package = "gsm.timez")
)
lReporting <- gsm.core::RunWorkflows(
  reporting_wf,
  c(lMapped, list(lAnalyzed = lAnalyzed, lWorkflows = metrics_wf))
)
```

## Generate HTML Report

The final step runs the `4_modules` workflows to generate an HTML report. This step
uses `gsm.kri::Report_KRI` to render the report template.

```{r run-modules, message=FALSE, results='hide'}
module_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/4_modules", package = "gsm.timez")
)
lReport <- gsm.core::RunWorkflows(
  module_wf,
  lReporting
)
```

```{r move-for-docs, include=FALSE}
# Move report to docs/articles/ during pkgdown builds
dest_dir <- file.path("..", "docs", "articles")
if (dir.exists(dest_dir)) {
  file.rename(lReport$Module_Timeline, file.path(dest_dir, "kri_report.html"))
}
```

The workflow generates `kri_report.html`. See the [Sample Report](kri_report.html) for an example output.
