---
title: "GSM YAML Workflow: HTML Report"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GSM YAML Workflow: HTML Report}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r vignette-setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(gsm.timez)
library(gsm.core)
library(gsm.mapping)
```

This vignette walks through the end-to-end GSM YAML workflow to generate an
HTML report for the timeline KRI using the funnel-based z-score method. It
follows the standard four-phase GSM pipeline: mapping raw data into a
standardised layer, running the metric analysis, assembling the reporting
data, and rendering the HTML output. The result is the
[Sample Report](Report_TimeZFunnel.html) linked in the navbar.

## Prepare Source Data

```{r prepare-source, message=FALSE}
lSource <- list(
  Source_STUDY = clindata::ctms_study,
  Source_SITE = clindata::ctms_site,
  Source_SUBJ = clindata::rawplus_dm,
  Source_AE = clindata::rawplus_ae,
  # gsm.mapping expects a studyid column, but clindata::edc_data_pages uses
  # protocolname instead. DATAENT itself is not used by gsm.timez but is
  # required by gsm.mapping's AE workflow.
  Source_DATAENT = clindata::edc_data_pages %>%
    dplyr::mutate(studyid = protocolname),
  Source_VISIT = clindata::rawplus_visdt
)
```

## Step 1 — Map Raw Data

Use `Ingest()` to standardise column names based on the workflow specs, then
run the mapping workflows to produce the `Mapped_*` data layer.

```{r run-mapping, message=FALSE, results='hide'}
mapping_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/1_mappings", package = "gsm.mapping"),
  strNames = c("SUBJ", "AE", "VISIT", "SITE", "STUDY", "COUNTRY")
)
mapping_spec <- gsm.mapping::CombineSpecs(mapping_wf)
lRaw <- gsm.mapping::Ingest(lSource, mapping_spec)
lMapped <- gsm.core::RunWorkflows(mapping_wf, lRaw)
```

```{r show-mapped}
names(lMapped)
```

## Step 2 — Run Analysis

Run the `kri_TimeZFunnel` metric workflow to produce the analysis results.

```{r run-metrics, message=FALSE, results='hide'}
metrics_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/2_metrics", package = "gsm.timez"),
  strNames = "kri_TimeZFunnel",
  bExact = TRUE
)
lAnalyzed <- gsm.core::RunWorkflows(metrics_wf, lMapped)
```

```{r show-analyzed}
names(lAnalyzed)
```

## Step 3 — Prepare Reporting Data

The `3_reporting` workflows stack and reshape the analysis results into the
flat data frames expected by the report template.

```{r run-reporting, message=FALSE, results='hide'}
reporting_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/3_reporting", package = "gsm.timez")
)
lReporting <- gsm.core::RunWorkflows(
  reporting_wf,
  c(lMapped, list(lAnalyzed = lAnalyzed, lWorkflows = metrics_wf))
)
```

```{r show-reporting}
names(lReporting)
```

## Step 4 — Generate HTML Report

The `4_modules` workflow creates all charts and renders the HTML report using
`gsm.timez`'s custom template (`inst/report/Report_TimeZFunnel.Rmd`).

```{r run-modules, message=FALSE, results='hide'}
module_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/4_modules", package = "gsm.timez"),
  strNames = "Report_TimeZFunnel",
  bExact = TRUE
)
lReport <- gsm.core::RunWorkflows(module_wf, lReporting)
```

```{r show-report-path}
lReport$Module_Report_TimeZFunnel
```

The workflow generates `Report_TimeZFunnel.html`. See the [Sample Report](Report_TimeZFunnel.html) for an example output.

```{r move-for-docs, include=FALSE}
# Move report to docs/articles/ during pkgdown builds
dest_dir <- file.path("..", "docs", "articles")
report_path <- lReport$Module_Report_TimeZFunnel
if (!is.null(report_path) && file.exists(report_path) && dir.exists(dest_dir)) {
  file.rename(report_path, file.path(dest_dir, "Report_TimeZFunnel.html"))
}
```
