---
title: "GSM Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GSM Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates two approaches to using `gsm.timez` for cumulative 
timeline analysis: direct function calls (scripted) and YAML workflow execution.

## Setup

```{r load-packages, message = FALSE}
library(gsm.timez)
library(gsm.core)
library(gsm.mapping)
library(dplyr)
```

## Prepare Mapped Data

Both approaches require mapped data from `gsm.mapping`. We'll use sample data 
from the `clindata` package.

```{r prepare-mapped-data, message = FALSE, results = 'hide'}
lRaw <- list(
  Raw_SUBJ = clindata::rawplus_dm,
  Raw_AE = clindata::rawplus_ae,
  Raw_VISIT = clindata::rawplus_visdt
)

all_mapping_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/1_mappings", package = "gsm.mapping")
)

mapping_wf <- all_mapping_wf[c("AE", "SUBJ", "VISIT")]
spec <- gsm.mapping::CombineSpecs(mapping_wf)
lIngest <- gsm.mapping::Ingest(lRaw, spec)
lMapped <- gsm.core::RunWorkflows(mapping_wf, lIngest)
```

```{r show-mapped-data}
cat("Mapped data available:\n")
cat("- Mapped_SUBJ:", nrow(lMapped$Mapped_SUBJ), "rows\n")
cat("- Mapped_AE:", nrow(lMapped$Mapped_AE), "rows\n")
cat("- Mapped_VISIT:", nrow(lMapped$Mapped_VISIT), "rows\n")
```

***

## Section 1: Scripted Approach

The scripted approach uses direct function calls without YAML configuration.
This is useful for quick analysis, prototyping, or when you need more control
over intermediate steps.

### Step 1: Generate Timeline

```{r scripted-timeline}
dfTimeline_scripted <- Timeline(
  dfSubjects = lMapped$Mapped_SUBJ,
  dfNumerator = lMapped$Mapped_AE,
  dfDenominator = lMapped$Mapped_VISIT,
  strSubjectCol = "subjid",
  strGroupCol = "invid",
  strNumeratorDateCol = "aest_dt",
  strDenominatorDateCol = "visit_dt"
)

cat("Timeline output:", nrow(dfTimeline_scripted), "rows\n")
knitr::kable(head(dfTimeline_scripted, 10))
```

### Step 2: Calculate Z-Scores

```{r scripted-zscore}
dfAnalyzed_scripted <- TimeZScore(dfTimeline_scripted)

cat("Analyzed output:", nrow(dfAnalyzed_scripted), "rows\n")
cat("New columns added: Metric, Score\n")
knitr::kable(head(dfAnalyzed_scripted, 10))
```

### Step 3: Flag Outliers

```{r scripted-flag}
dfFlagged_scripted <- gsm.timez::Flag(dfAnalyzed_scripted, vThreshold = c(-2, -1, 1, 2))

cat("Flagged output:", nrow(dfFlagged_scripted), "rows\n")
cat("New column added: Flag\n")
knitr::kable(head(dfFlagged_scripted, 10))
```

### Scripted Approach: Chained Pipeline

You can also chain the functions together using pipes:
```{r scripted-chained}
dfFlagged_chained <- Timeline(
  dfSubjects = lMapped$Mapped_SUBJ,
  dfNumerator = lMapped$Mapped_AE,
  dfDenominator = lMapped$Mapped_VISIT,
  strSubjectCol = "subjid",
  strGroupCol = "invid",
  strNumeratorDateCol = "aest_dt",
  strDenominatorDateCol = "visit_dt"
) %>%
  TimeZScore() %>%
  gsm.timez::Flag(vThreshold = c(-2, -1, 1, 2))

cat("Chained pipeline output:", nrow(dfFlagged_chained), "rows\n")
```

***

## Section 2: YAML Workflow Approach

The YAML workflow approach uses the GSM workflow system, which is designed for
enterprise-scale Risk Based Quality Monitoring (RBQM). This provides:

- Declarative, reproducible analysis definitions
- Integration with the broader GSM ecosystem
- Standardized reporting pipelines

### Load and Run the Workflow

```{r yaml-workflow, message = FALSE, results = 'hide'}
timez_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/2_metrics", package = "gsm.timez")
)

cat("Available workflows:", names(timez_wf), "\n")
```

```{r yaml-run}
lAnalysis <- gsm.core::RunWorkflows(timez_wf, lMapped)
```

### Inspect Workflow Output

The workflow returns a list containing all intermediate and final results:

```{r yaml-output}
cat("Workflow output components:\n")
names(lAnalysis$Analysis_kri_timez)
```

```{r yaml-results}
dfFlagged_yaml <- lAnalysis$Analysis_kri_timez$Flagged

cat("Flagged output from YAML workflow:", nrow(dfFlagged_yaml), "rows\n")
knitr::kable(head(dfFlagged_yaml, 10))
```

***

## Comparison: Scripted vs YAML

Both approaches produce equivalent results:

```{r compare-outputs}
scripted_result <- dfFlagged_scripted %>%
  select(GroupID, GroupLevel, NMonth, Numerator, Denominator, Metric, Score, Flag) %>%
  arrange(GroupID, NMonth)

yaml_result <- dfFlagged_yaml %>%
  select(GroupID, GroupLevel, NMonth, Numerator, Denominator, Metric, Score, Flag) %>%
  arrange(GroupID, NMonth)

cat("Results are equivalent:", all.equal(scripted_result, yaml_result), "\n")
```

***

## Visualization

Both outputs can be visualized using the `Visualize()` function:

```{r visualize, fig.width = 8, fig.height = 5}
Visualize(dfFlagged_scripted)
```

***

## Summary

| Approach | Best For | Pros | Cons |
|----------|----------|------|------|
| **Scripted** | Quick analysis, prototyping | Simple, flexible, easy to debug | Manual parameter management |
| **YAML Workflow** | Enterprise RBQM, reproducibility | Declarative, integrates with GSM pipeline | Requires YAML configuration |

Choose the approach that best fits your use case. For integration with the 
full GSM ecosystem (including reporting and multi-study analysis), use the 
YAML workflow approach.
