---
title: "HTML Report Generation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HTML Report Generation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates how to generate interactive HTML reports using 
`gsm.timez`. The package provides plotly-based widgets and a complete report 
generation function following GSM conventions.

## Setup

```{r setup, message=FALSE}
library(gsm.timez)
library(dplyr)
```

## Prepare Data

First, generate flagged timeline data using the standard pipeline:

```{r data-prep}
dfSubjects <- clindata::rawplus_dm
dfNumerator <- clindata::rawplus_ae
dfDenominator <- clindata::rawplus_visdt %>%
  dplyr::mutate(visit_dt = as.Date(visit_dt, "%Y-%m-%d"))

dfFlagged <- Timeline(
  dfSubjects = dfSubjects,
  dfNumerator = dfNumerator,
  dfDenominator = dfDenominator,
  strGroupCol = "siteid",
  strSubjectCol = "subjid",
  strNumeratorDateCol = "aest_dt",
  strDenominatorDateCol = "visit_dt"
) %>%
  TimeZScore() %>%
  Flag()

knitr::kable(head(dfFlagged))
```

## Interactive Widgets

### Widget_Timeline

The `Widget_Timeline()` function creates an interactive line chart showing 
metrics over time for each group. Features include:

- Hover tooltips with detailed information
- Zoom and pan capabilities
- Points colored by flag status
- Click to highlight individual groups

```{r widget-timeline, fig.width=8, fig.height=6}
Widget_Timeline(dfFlagged)
```

The widget wraps the static `Visualize()` function and converts it to an 
interactive plotly chart using `ggplotly()`.

### Widget_TimelineBoxplot

The `Widget_TimelineBoxplot()` function creates an interactive boxplot showing 
the distribution of metric values for each month:

```{r widget-boxplot, fig.width=8, fig.height=6}
Widget_TimelineBoxplot(dfFlagged)
```

#### Filter to Specific Months

Use the `nMonths` parameter to focus on a subset of months:

```{r widget-boxplot-filtered, fig.width=8, fig.height=6}
Widget_TimelineBoxplot(dfFlagged, nMonths = 1:12)
```

## Generating Chart Lists

The `MakeCharts_Timeline()` function generates a list of both widgets, useful 
for batch processing or integration with custom reports:

```{r make-charts}
lCharts <- MakeCharts_Timeline(dfFlagged)
names(lCharts)
```

You can use a custom MetricID to distinguish charts when combining with other 
GSM metrics:

```{r make-charts-custom}
lCharts <- MakeCharts_Timeline(dfFlagged, strMetricID = "AE_Rate")
names(lCharts)
```

## Complete HTML Report

The `Report_Timeline()` function generates a complete HTML report with:

- Summary statistics
- Flag distribution chart
- Interactive timeline visualization
- Interactive boxplot
- Table of currently flagged groups
- Data preview

### Basic Usage

```{r report-basic, eval=FALSE}
Report_Timeline(
  dfFlagged = dfFlagged,
  strOutputFile = "timeline_report.html"
)
```

### With Study Information

```{r report-study, eval=FALSE}
Report_Timeline(
  dfFlagged = dfFlagged,
  strStudyID = "STUDY-001",
  strGroupLevel = "Site",
  strOutputDir = "reports",
  strOutputFile = "site_timeline_report.html"
)
```

### With Pre-generated Charts

For more control, you can pre-generate charts and pass them to the report:

```{r report-charts, eval=FALSE}
lCharts <- MakeCharts_Timeline(dfFlagged)

Report_Timeline(
  dfFlagged = dfFlagged,
  lCharts = lCharts,
  strStudyID = "STUDY-001",
  strOutputFile = "custom_report.html"
)
```

## Report Contents

The generated HTML report includes the following sections:

| Section | Description |
|---------|-------------|
| Overview | Study ID, report date, group level |
| Summary Statistics | Total groups, months, observations, flag counts |
| Flag Distribution | Bar chart of flag categories |
| Timeline Chart | Interactive line plot of all groups over time |
| Boxplot | Distribution of metrics by month |
| Flagged Groups | Sortable table of currently flagged groups |
| Data Preview | First 100 rows of the timeline data |

## Workflow Integration

The reporting functionality is also available as a YAML workflow step in 
`inst/workflow/3_reporting/Report.yaml`:

```{r workflow-report, eval=FALSE}
library(gsm.core)

# After running mapping and metric workflows...
report_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/3_reporting", package = "gsm.timez")
)

lReport <- gsm.core::RunWorkflows(report_wf["Report"], lData)
```

## Color Scheme

All visualizations use a consistent diverging color scheme for flags:
| Flag | Label | Color |
|------|-------|-------|
| -2 | Red Low | #D73027 |
| -1 | Amber Low | #FC8D59 |
| 0 | Normal | #999999 |
| 1 | Amber High | #91BFDB |
| 2 | Red High | #4575B4 |

## Dependencies

The reporting functions require:

- `plotly` - Interactive visualizations
- `htmltools` - HTML generation
- `rmarkdown` - Report rendering
- `DT` (optional) - Interactive tables in reports

These are automatically installed when you install `gsm.timez`.
