---
title: "Using YAML Workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using YAML Workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette demonstrates how to use `gsm.timez` with the GSM YAML workflow 
system, following the pattern used by other IMPALA packages like `gsm.studykri`.

## Overview

The GSM workflow system allows you to define analysis pipelines using YAML 
configuration files. This approach provides:

- Reproducible, declarative analysis definitions
- Integration with the broader GSM ecosystem
- Standardized data mapping through `gsm.mapping`

Workflow files for `gsm.timez` are located in:

- `inst/workflow/2_metrics/` - Analysis workflows (Timeline, TimeZScore, Flag)
- `inst/workflow/3_visualization/` - Visualization workflows (Visualize)
- `inst/workflow/3_reporting/` - Reporting workflows (Report)

## Setup

```{r setup, message=FALSE}
library(gsm.timez)
library(gsm.core)
library(gsm.mapping)
library(dplyr)
```

## Step 1: Load Raw Data

Load raw clinical data into a named list with `Raw_` prefixes. This example 
uses sample data from the `clindata` package:

```{r load-raw-data}
lRaw <- list(
  Raw_SUBJ = clindata::rawplus_dm,
  Raw_AE = clindata::rawplus_ae,
  Raw_VISIT = clindata::rawplus_visdt
)

cat("Raw data loaded:\n")
cat("- Raw_SUBJ:", nrow(lRaw$Raw_SUBJ), "rows\n")
cat("- Raw_AE:", nrow(lRaw$Raw_AE), "rows\n")
cat("- Raw_VISIT:", nrow(lRaw$Raw_VISIT), "rows\n")
```

## Step 2: Run Mapping Workflows

Use `gsm.mapping` to transform raw data into standardized mapped domains. 
We load all available mappings but filter to only the domains needed for 
the Timeline analysis:

```{r run-mappings, message=FALSE, results='hide'}
all_mapping_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/1_mappings", package = "gsm.mapping")
)

mapping_wf <- all_mapping_wf[c("AE", "SUBJ", "VISIT")]

spec <- gsm.mapping::CombineSpecs(mapping_wf)
lIngest <- gsm.mapping::Ingest(lRaw, spec)
lMapped <- gsm.core::RunWorkflows(mapping_wf, lIngest)
```

The mapped data now contains standardized domain data:

```{r show-mapped}
names(lMapped)
```

## Step 3: Run Timeline Workflow

Load the `gsm.timez` metric workflows and run the Timeline workflow:

```{r run-timeline, message=FALSE, results='hide'}
metrics_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/2_metrics", package = "gsm.timez")
)

cat("Available workflows:", names(metrics_wf), "\n")

timeline_wf <- metrics_wf["Timeline"]
lTimeline <- gsm.core::RunWorkflows(timeline_wf, lMapped)
```

## Step 4: Access Timeline Results

The Timeline workflow produces site-level cumulative event data:

```{r timeline-results}
dfTimeline <- lTimeline$Analysis_Timeline
knitr::kable(head(dfTimeline, 10))
```

## Step 5: Run TimeZScore Workflow (Chained)
 
The TimeZScore workflow can be chained after Timeline. It requires the 
`Analysis_Timeline` output from the previous step:

```{r run-timezscore, message=FALSE, results='hide'}
lData <- c(lMapped, lTimeline)

timezscore_wf <- metrics_wf["TimeZScore"]
lResults <- gsm.core::RunWorkflows(timezscore_wf, lData)
```

```{r timezscore-results}
dfTimeZScore <- lResults$Analysis_TimeZScore
knitr::kable(head(dfTimeZScore, 10))
```

## Alternative: Direct Function Call

You can also call `TimeZScore()` directly on the Timeline output instead of 
using the workflow:

```{r direct-call}
dfTimeZScore_direct <- gsm.timez::TimeZScore(dfTimeline)
knitr::kable(head(dfTimeZScore_direct))
```

## Step 6: Run Flag Workflow

The Flag workflow adds statistical outlier flags based on z-score thresholds:

```{r run-flag, message=FALSE, results='hide'}
lData <- c(lMapped, lTimeline, lResults)

flag_wf <- metrics_wf["Flag"]
lFlagged <- gsm.core::RunWorkflows(flag_wf, lData)
```

```{r flag-results}
dfFlagged <- lFlagged$Analysis_Flag
knitr::kable(head(dfFlagged, 10))
```

## Step 7: Run Visualization Workflow

Load the visualization workflows and create a chart:
  
```{r run-visualize, message=FALSE, results='hide'}
viz_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/3_visualization", package = "gsm.timez")
)

cat("Available visualization workflows:", names(viz_wf), "\n")

lData <- c(lData, lFlagged)
lViz <- gsm.core::RunWorkflows(viz_wf["Visualize"], lData)
```

```{r show-chart, fig.width=8, fig.height=5}
lViz$Visualization_Visualize
```

## Step 8: Generate HTML Report (Optional)

Reporting workflows are available in `inst/workflow/3_reporting/`. See the 
[HTML Report Generation](HTMLReporting.html) vignette for details.

## Complete Pipeline

Here's the complete workflow from raw data to visualization using only workflows:

```{r complete-pipeline, message=FALSE, results='hide'}
lRaw <- list(
  Raw_SUBJ = clindata::rawplus_dm,
  Raw_AE = clindata::rawplus_ae,
  Raw_VISIT = clindata::rawplus_visdt
)

all_mapping_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/1_mappings", package = "gsm.mapping")
)
mapping_wf <- all_mapping_wf[c("AE", "SUBJ", "VISIT")]
spec <- gsm.mapping::CombineSpecs(mapping_wf)
lIngest <- gsm.mapping::Ingest(lRaw, spec)
lMapped <- gsm.core::RunWorkflows(mapping_wf, lIngest)

metrics_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/2_metrics", package = "gsm.timez")
)
viz_wf <- gsm.core::MakeWorkflowList(
  strPath = system.file("workflow/3_visualization", package = "gsm.timez")
)

lTimeline <- gsm.core::RunWorkflows(metrics_wf["Timeline"], lMapped)
lData <- c(lMapped, lTimeline)

lTimeZScore <- gsm.core::RunWorkflows(metrics_wf["TimeZScore"], lData)
lData <- c(lData, lTimeZScore)

lFlagged <- gsm.core::RunWorkflows(metrics_wf["Flag"], lData)
lData <- c(lData, lFlagged)

lViz <- gsm.core::RunWorkflows(viz_wf["Visualize"], lData)
```

```{r complete-results, fig.width=8, fig.height=5}
lViz$Visualization_Visualize
```

## Alternative: Direct Function Calls

You can also mix workflow execution with direct function calls:

```{r direct-visualization, fig.width=8, fig.height=5}
dfFlagged <- lTimeline$Analysis_Timeline %>%
  gsm.timez::TimeZScore() %>%
  gsm.timez::Flag()

gsm.timez::Visualize(dfFlagged)
```

## Workflow YAML Structure

The Timeline workflow YAML (`inst/workflow/2_metrics/Timeline.yaml`) defines:

- **meta**: Workflow metadata (ID, description, type)
- **spec**: Required input data frames and their column specifications
- **steps**: The function calls to execute

```yaml
meta:
  Type: Analysis
  ID: Timeline
  Description: Generate site-level timeline of cumulative events

spec:
  Mapped_SUBJ:
    subjid:
      type: character
    invid:
      type: character
  Mapped_AE:
    subjid:
      type: character
    aest_dt:
      type: Date
  Mapped_VISIT:
    subjid:
      type: character
    visit_dt:
      type: Date

steps:
  - output: Analysis_Timeline
    name: gsm.timez::Timeline
    params:
      dfSubjects: Mapped_SUBJ
      dfNumerator: Mapped_AE
      dfDenominator: Mapped_VISIT
      strGroupCol: invid
      strSubjectCol: subjid
      strNumeratorDateCol: aest_dt
      strDenominatorDateCol: visit_dt
```
